---
title: "3 Data visualisation"
output: html_notebook
---

## first steps

```{r}
xfun::pkg_attach2(c("tidyverse"))
# You can also call functions or datasets from a package::function() pr package::dataset
# if the package is installed
```

### Do cars with big engines use more fuel than cars with small engines?

We are going to use the mpg data frame found in the ggplot2 package, i.e. ggplot2::mpg()
Among the variables in mpg are:

* displ, a car’s engine size, in litres.

* hwy, a car’s fuel efficiency on the highway, in miles per gallon (mpg). A car with a low fuel efficiency consumes more fuel than a car with a high fuel efficiency when they travel the same distance.

To learn more about mpg, open its help page by running ?mpg.

```{r}

mpg
?mpg
```

To plot mpg, run this code to put displ on the x-axis and hwy on the y-axis:

```{r}
ggplot(data = mpg)+
  geom_point(mapping= aes(x=displ, y= hwy))
```

With ggplot2, you begin a plot with the function ggplot(). ggplot() creates a coordinate system that you can add layers to. The first argument of ggplot() is the dataset to use in the graph. So ggplot(data = mpg) creates an empty graph, but it’s not very interesting so I’m not going to show it here.

You complete your graph by adding one or more layers to ggplot(). The function geom_point() adds a layer of points to your plot, which creates a scatterplot. ggplot2 comes with many geom functions that each add a different type of layer to a plot. You’ll learn a whole bunch of them throughout this chapter.

Each geom function in ggplot2 takes a mapping argument. This defines how variables in your dataset are mapped to visual properties. The mapping argument is always paired with aes(), and the x and y arguments of aes() specify which variables to map to the x and y axes. ggplot2 looks for the mapped variables in the data argument, in this case, mpg.

### A graphing template
To make a graph, replace the bracketed sections in the code below with a dataset, a geom function, or a collection of mappings.

ggplot(data = <DATA>) +  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

### 3.2.4. Exercises

1. I see an empty graph area

2. 243 rows. 11 columns.

3. drv is the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

4. the plot is below

5. the varibales are categorical and hence show no correlation or concentration at intersection


```{r}
ggplot(data = mpg)
mpg
?mpg

ggplot(data = mpg) + 
  geom_point(aes(x=hwy, y =  cyl))

ggplot(data = mpg) +
  geom_point(aes(x=class, y= drv))
```

## Aesthetic mappings

An aesthetic is a visual property of the objects in your plot. Aesthetics include things like the size, the shape, or the color of your points. You can display a point in different ways by changing the values of its aesthetic properties. Since we already use the word “value” to describe data, let’s use the word “level” to describe aesthetic properties.

You can convey information about your data by mapping the aesthetics in your plot to the variables in your dataset. For example, you can map the colors of your points to the class variable to reveal the class of each car.

```{r}
library(tidyverse)
ggplot(data = mpg)+
  geom_point(aes(x= displ, y= hwy, colour = class)) #You can use color or colour

##we can also do size though it's not advisable for a discrete variable
##We get a warning here, because mapping an unordered variable (class) to an ordered aesthetic (size) is not a good idea.

ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y = hwy, size = class))

#Or we could have mapped class to the alpha aesthetic, which controls the transparency of the points, or to the shape aesthetic, which controls the shape of the points

ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y = hwy, alpha = class))

ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))

```

*On shape*

What happened to the SUVs? ggplot2 will only use six shapes at a time. By default, additional groups will go unplotted when you use the shape aesthetic.

For each aesthetic, you use aes() to associate the name of the aesthetic with a variable to display.

You can also set the aesthetic properties of your geom manually. For example, we can make all of the points in our plot blue:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), colour = "red")

```

To set an aesthetic manually, set the aesthetic by name as an argument of your geom function; i.e. it goes outside of aes(). You’ll need to pick a level that makes sense for that aesthetic.

* The name of a color as a character string.

* The size of a point in mm.

* The shape of a point as a number from 0 to 14, then duplicates thereafter to 24.

### Exercise 3.3.1

1. The manual aesthetic mapping is inside the aes() function

2. 6 are categorical (manufacturer, model, trans, drv, fl, class) and 5 are Continuos (displ, year, cyl, cty, hwy). This information is below the variable names in the dataset display.


```{r}
?mpg
mpg
```

3. size and color scale, while shape returns error

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = cty))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = cty))

try(ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = cty)))
```


4. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = hwy, color = hwy))
```

5. Stroke makes the points big and overlapping

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy, stroke = cty))

```

6. A condition in aes returns true or false mapping

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))

```

### 3.5 Facets

Facets helps us add a discrete variable in addition to other variables in aes()

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2) # for one variable

## use facet_grid for 2 variables
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)

```

3.5.1 Exercises

1. For a continuous variable

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(.~ cty)
```

2. There are no data points (no observations) for drv and cyl for such a combination of points

3. What plots does the following code make? What does . do?. "." ignores dimensions while plotting .~ ignores y axis and plots values by x axis, while  ~. ignores x axis

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

4. more visibility, less overview

5. nrow and ncol gives values for rows or columns for the facet_wrap()

6. To maintain the landscape view of the overall image


### 3.6 Geometric objects

*A geom is the geometrical object that a plot uses to represent data.*

e.g. geom smooth

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )

ggplot(data = mpg) +  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv)) +
  geom_point(aes(x = displ, y = hwy, color = drv))

```

we can also pass more geometric objects (geoms) better

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(aes(color= class)) + geom_smooth()
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)

```

> 1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

geom_line(), geom_boxplot(), gemo_histogram(), geom_area()

> 2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions

the result has colors over drv categories and standard errors have been ignored for the smooth lines over drv categories.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```


> 3. What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

It hides the legend box of the chart for additional variable(S). If removed the legend shows. To hide multiple legends on the many charts.

> 4. What does the se argument to geom_smooth() do?

it adds the upper and lower standard error curves/bands to the smoothing lines

> 5. Will these two graphs look different? Why/why not?

NO. The first graph places repetitive attributes of data and mappings

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

> 6. Recreate the R code necessary to generate the following graphs.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy))+ geom_point() +geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv))+ geom_point() +geom_smooth(se = FALSE) 
# group variable is used only by smooth, so we can assign it smooth mapping option alone

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(group = drv), se = FALSE) +
  geom_point()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth(se = FALSE)

ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy), se = FALSE)

# We could customise those not common to each geometric object
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = drv)) +
  geom_smooth(se = FALSE)


ggplot(data = mpg)+ 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv), se = FALSE)
## Customise as well

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = drv)) +
  geom_smooth(aes(linetype = drv), se = FALSE)

##Assigning colour of the border of object and color of the fill
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(size = 4, color = "white") +
  geom_point(aes(colour = drv))
```


## 3.7 Statistical transformations

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
  
```

Many graphs, like scatterplots, plot the raw values of your dataset. Other graphs, like bar charts, calculate new values to plot:

* bar charts, histograms, and frequency polygons bin your data and then plot bin counts, the number of points that fall in each bin.

* smoothers fit a model to your data and then plot predictions from the model.

* boxplots compute a robust summary of the distribution and then display a specially formatted box.

The algorithm used to calculate new values for a graph is called a stat, short for statistical transformation.

?geom_bar shows that the default value for stat is “count”, which means that geom_bar() uses stat_count()

You can generally use geoms and stats interchangeably. For example, you can recreate the previous plot using stat_count() instead of geom_bar():

```{r}
ggplot(data = diamonds) +
  stat_count(mapping = aes(x = cut))
```

This works because every geom has a default stat; and every stat has a default geom. This means that you can typically use geoms without worrying about the underlying statistical transformation. There are three reasons you might need to use a stat explicitly:

1. You might want to override the default stat

```{r}
data2 <- tribble(
  ~cut, ~freq,
  "Fair", 1610,
  "Good", 12000,
  "Ideal", 8000
)

ggplot(data = data2) +
  geom_bar(mapping = aes(x = cut, y = freq), stat =  "identity")

```

2. You might want to override the default mapping from transformed variables to aesthetics. For example, you might want to display a bar chart of proportion, rather than count:

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1)) ## you can replace stat(prop) with ..prop..

```
See help for *“computed variables”*

3. You might want to draw greater attention to the statistical transformation in your code. For example, you might use stat_summary()

```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median ##default is mean
  )

ggplot(data = diamonds) + 
  geom_pointrange(mapping = aes(x = cut, y = depth), 
                                          stat = "summary",  fun.min = min, fun.max = max, fun = median)

```


#### 3.7.1 Exercises

> 1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

geom_pointrange()

> 2. What does geom_col() do? How is it different to geom_bar()?

plots x as bars of y. Both x and y should be in the data, so it only uses stat_identity(). Geom_bar() can use one variable and derive y from it

```{r}
ggplot(data = data2) +
  geom_col(mapping = aes(x = cut, y = freq))

```
The geom_col() function has different default stat than geom_bar(). The default stat of geom_col() is stat_identity(), which leaves the data as is. The geom_col() function expects that the data contains x values and y values which represent the bar height.

> 3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

Many geoms have stat_identity() as the default stat.

see geom_stat.ods file for the groupings

> 4. What variables does stat_smooth() compute? What parameters control its behavior?

Predicted y and prediction ranges given a confidence interval. 

The function stat_smooth() calculates the following variables:

* y: predicted value
* ymin: lower value of the confidence interval
* ymax: upper value of the confidence interval
* se: standard error

The “Computed Variables” section of the stat_smooth() documentation contains these variables.

The parameters that control the behavior of stat_smooth() include:

* method: This is the method used to compute the smoothing line. If NULL, a default method is used based on the sample size: stats::loess() when there are less than 1,000 observations in a group, and mgcv::gam() with formula = y ~ s(x, bs = "CS") otherwise. Alternatively, the user can provide a character vector with a function name, e.g. "lm", "loess", or a function, e.g. MASS::rlm.

* formula: When providing a custom method argument, the formula to use. The default is y ~ x. For example, to use the line implied by lm(y ~ x + I(x ^ 2) + I(x ^ 3)), use method = "lm" or method = lm and formula = y ~ x + I(x ^ 2) + I(x ^ 3).

* method.arg(): Arguments other than than the formula, which is already specified in the formula argument, to pass to the function inmethod`.

* se: If TRUE, display standard error bands, if FALSE only display the line.

* na.rm: If FALSE, missing values are removed with a warning, if TRUE the are silently removed. They default is FALSE in order to make debugging easier. If missing values are known to be in the data, then can be ignored, but if missing values are not anticipated this warning can help catch errors.

> 5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

This ensures the proportions sum to 1 as a whole. The group was not set.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```


If group = 1 is not included, then all the bars in the plot will have the same height, a height of 1. The function geom_bar() assumes that the groups are equal to the x values since the stat computes the counts within the group.

The problem with these two plots is that the proportions are calculated within the groups.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..count.. / sum(..count..), fill = color))
```

<br>

## 3.8 Position adjustments














