---
title: "3 Data visualisation"
output: html_notebook
---

## first steps

```{r}
xfun::pkg_attach2(c("tidyverse"))
# You can also call functions or datasets from a package::function() pr package::dataset
# if the package is installed
```

### Do cars with big engines use more fuel than cars with small engines?

We are going to use the mpg data frame found in the ggplot2 package, i.e. ggplot2::mpg()
Among the variables in mpg are:

* displ, a car’s engine size, in litres.

* hwy, a car’s fuel efficiency on the highway, in miles per gallon (mpg). A car with a low fuel efficiency consumes more fuel than a car with a high fuel efficiency when they travel the same distance.

To learn more about mpg, open its help page by running ?mpg.

```{r}

mpg
?mpg
```

To plot mpg, run this code to put displ on the x-axis and hwy on the y-axis:

```{r}
ggplot(data = mpg)+
  geom_point(mapping= aes(x=displ, y= hwy))
```

With ggplot2, you begin a plot with the function ggplot(). ggplot() creates a coordinate system that you can add layers to. The first argument of ggplot() is the dataset to use in the graph. So ggplot(data = mpg) creates an empty graph, but it’s not very interesting so I’m not going to show it here.

You complete your graph by adding one or more layers to ggplot(). The function geom_point() adds a layer of points to your plot, which creates a scatterplot. ggplot2 comes with many geom functions that each add a different type of layer to a plot. You’ll learn a whole bunch of them throughout this chapter.

Each geom function in ggplot2 takes a mapping argument. This defines how variables in your dataset are mapped to visual properties. The mapping argument is always paired with aes(), and the x and y arguments of aes() specify which variables to map to the x and y axes. ggplot2 looks for the mapped variables in the data argument, in this case, mpg.

### A graphing template
To make a graph, replace the bracketed sections in the code below with a dataset, a geom function, or a collection of mappings.

ggplot(data = <DATA>) +  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

### 3.2.4. Exercises

1. I see an empty graph area

2. 243 rows. 11 columns.

3. drv is the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

4. the plot is below

5. the varibales are categorical and hence show no correlation or concentration at intersection


```{r}
ggplot(data = mpg)
mpg
?mpg

ggplot(data = mpg) + 
  geom_point(aes(x=hwy, y =  cyl))

ggplot(data = mpg) +
  geom_point(aes(x=class, y= drv))
```

## Aesthetic mappings

An aesthetic is a visual property of the objects in your plot. Aesthetics include things like the size, the shape, or the color of your points. You can display a point in different ways by changing the values of its aesthetic properties. Since we already use the word “value” to describe data, let’s use the word “level” to describe aesthetic properties.

You can convey information about your data by mapping the aesthetics in your plot to the variables in your dataset. For example, you can map the colors of your points to the class variable to reveal the class of each car.

```{r}
library(tidyverse)
ggplot(data = mpg)+
  geom_point(aes(x= displ, y= hwy, colour = class)) #You can use color or colour

##we can also do size though it's not advisable for a discrete variable
##We get a warning here, because mapping an unordered variable (class) to an ordered aesthetic (size) is not a good idea.

ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y = hwy, size = class))

#Or we could have mapped class to the alpha aesthetic, which controls the transparency of the points, or to the shape aesthetic, which controls the shape of the points

ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y = hwy, alpha = class))

ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))

```

*On shape*

What happened to the SUVs? ggplot2 will only use six shapes at a time. By default, additional groups will go unplotted when you use the shape aesthetic.

For each aesthetic, you use aes() to associate the name of the aesthetic with a variable to display.

You can also set the aesthetic properties of your geom manually. For example, we can make all of the points in our plot blue:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), colour = "red")

```

To set an aesthetic manually, set the aesthetic by name as an argument of your geom function; i.e. it goes outside of aes(). You’ll need to pick a level that makes sense for that aesthetic.

* The name of a color as a character string.

* The size of a point in mm.

* The shape of a point as a number from 0 to 14, then duplicates thereafter to 24.

### Exercise 3.3.1

1. The manual aesthetic mapping is inside the aes() function

2. 6 are categorical (manufacturer, model, trans, drv, fl, class) and 5 are Continuos (displ, year, cyl, cty, hwy). This information is below the variable names in the dataset display.


```{r}
?mpg
mpg
```

3. size and color scale, while shape returns error

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = cty))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = cty))

try(ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = cty)))
```


4. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = hwy, y = hwy, color = hwy))
```

5. Stroke makes the points big and overlapping

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy, stroke = cty))

```

6. A condition in aes returns true or false mapping

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))

```

## 3.5 Facets

Facets helps us add a discrete variable in addition to other variables in aes()

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2) # for one variable

## use facet_grid for 2 variables
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)

```

3.5.1 Exercises

1. For a continuous variable

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(.~ cty)
```

2. There are no data points (no observations) for drv and cyl for such a combination of points

3. What plots does the following code make? What does . do?. "." ignores dimensions while plotting .~ ignores y axis and plots values by x axis, while  ~. ignores x axis

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

4. more visibility, less overview

5. nrow and ncol gives values for rows or columns for the facet_wrap()

6. To maintain the landscape view of the overall image


## 3.6 Geometric objects

*A geom is the geometrical object that a plot uses to represent data.*

e.g. geom smooth

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )

ggplot(data = mpg) +  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv)) +
  geom_point(aes(x = displ, y = hwy, color = drv))

```

we can also pass more geometric objects (geoms) better

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(aes(color= class)) + geom_smooth()
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)

```

> 1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

geom_line(), geom_boxplot(), gemo_histogram(), geom_area()

> 2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions

the result has colors over drv categories and standard errors have been ignored for the smooth lines over drv categories.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```


> 3. What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

It hides the legend box of the chart for additional variable(S). If removed the legend shows. To hide multiple legends on the many charts.

> 4. What does the se argument to geom_smooth() do?

it adds the upper and lower standard error curves/bands to the smoothing lines

> 5. Will these two graphs look different? Why/why not?

NO. The first graph places repetitive attributes of data and mappings

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

> 6. Recreate the R code necessary to generate the following graphs.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy))+ geom_point() +geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv))+ geom_point() +geom_smooth(se = FALSE) 
# group variable is used only by smooth, so we can assign it smooth mapping option alone

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(group = drv), se = FALSE) +
  geom_point()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth(se = FALSE)

ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy), se = FALSE)

# We could customise those not common to each geometric object
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = drv)) +
  geom_smooth(se = FALSE)


ggplot(data = mpg)+ 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv), se = FALSE)
## Customise as well

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = drv)) +
  geom_smooth(aes(linetype = drv), se = FALSE)

##Assigning colour of the border of object and color of the fill
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(size = 4, color = "white") +
  geom_point(aes(colour = drv))
```


## 3.7 Statistical transformations

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
  
```

Many graphs, like scatterplots, plot the raw values of your dataset. Other graphs, like bar charts, calculate new values to plot:

* bar charts, histograms, and frequency polygons bin your data and then plot bin counts, the number of points that fall in each bin.

* smoothers fit a model to your data and then plot predictions from the model.

* boxplots compute a robust summary of the distribution and then display a specially formatted box.

The algorithm used to calculate new values for a graph is called a stat, short for statistical transformation.

?geom_bar shows that the default value for stat is “count”, which means that geom_bar() uses stat_count()

You can generally use geoms and stats interchangeably. For example, you can recreate the previous plot using stat_count() instead of geom_bar():

```{r}
ggplot(data = diamonds) +
  stat_count(mapping = aes(x = cut))
```

This works because every geom has a default stat; and every stat has a default geom. This means that you can typically use geoms without worrying about the underlying statistical transformation. There are three reasons you might need to use a stat explicitly:

1. You might want to override the default stat

```{r}
data2 <- tribble(
  ~cut, ~freq,
  "Fair", 1610,
  "Good", 12000,
  "Ideal", 8000
)

ggplot(data = data2) +
  geom_bar(mapping = aes(x = cut, y = freq), stat =  "identity")

```

2. You might want to override the default mapping from transformed variables to aesthetics. For example, you might want to display a bar chart of proportion, rather than count:

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1)) ## you can replace stat(prop) with ..prop..

```
See help for *“computed variables”*

3. You might want to draw greater attention to the statistical transformation in your code. For example, you might use stat_summary()

```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median ##default is mean
  )

ggplot(data = diamonds) + 
  geom_pointrange(mapping = aes(x = cut, y = depth), 
                                          stat = "summary",  fun.min = min, fun.max = max, fun = median)

```


#### 3.7.1 Exercises

> 1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

geom_pointrange()

> 2. What does geom_col() do? How is it different to geom_bar()?

plots x as bars of y. Both x and y should be in the data, so it only uses stat_identity(). Geom_bar() can use one variable and derive y from it

```{r}
ggplot(data = data2) +
  geom_col(mapping = aes(x = cut, y = freq))

```
The geom_col() function has different default stat than geom_bar(). The default stat of geom_col() is stat_identity(), which leaves the data as is. The geom_col() function expects that the data contains x values and y values which represent the bar height.

> 3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

Many geoms have stat_identity() as the default stat.

see geom_stat.ods file for the groupings

> 4. What variables does stat_smooth() compute? What parameters control its behavior?

Predicted y and prediction ranges given a confidence interval. 

The function stat_smooth() calculates the following variables:

* y: predicted value
* ymin: lower value of the confidence interval
* ymax: upper value of the confidence interval
* se: standard error

The “Computed Variables” section of the stat_smooth() documentation contains these variables.

The parameters that control the behavior of stat_smooth() include:

* method: This is the method used to compute the smoothing line. If NULL, a default method is used based on the sample size: stats::loess() when there are less than 1,000 observations in a group, and mgcv::gam() with formula = y ~ s(x, bs = "CS") otherwise. Alternatively, the user can provide a character vector with a function name, e.g. "lm", "loess", or a function, e.g. MASS::rlm.

* formula: When providing a custom method argument, the formula to use. The default is y ~ x. For example, to use the line implied by lm(y ~ x + I(x ^ 2) + I(x ^ 3)), use method = "lm" or method = lm and formula = y ~ x + I(x ^ 2) + I(x ^ 3).

* method.arg(): Arguments other than than the formula, which is already specified in the formula argument, to pass to the function inmethod`.

* se: If TRUE, display standard error bands, if FALSE only display the line.

* na.rm: If FALSE, missing values are removed with a warning, if TRUE the are silently removed. They default is FALSE in order to make debugging easier. If missing values are known to be in the data, then can be ignored, but if missing values are not anticipated this warning can help catch errors.

> 5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

This ensures the proportions sum to 1 as a whole. The group was not set.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```


If group = 1 is not included, then all the bars in the plot will have the same height, a height of 1. The function geom_bar() assumes that the groups are equal to the x values since the stat computes the counts within the group.

The problem with these two plots is that the proportions are calculated within the groups.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..count.. / sum(..count..), fill = color))
```

<br>

## 3.8 Position adjustments

You can colour a bar chart using either the colour aesthetic, or, more usefully, fill

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

Position adjustment can be achieved by using options "identity", "dodge" or "fill"

* position = "identity" will place each object as it falls in the context of the graph. This is not very useful for bars, because it overlaps them. To see that overlapping we either need to make the bars slightly transparent by setting alpha to a small value, or completely transparent by setting fill = NA

```{r}
ggplot(data=diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(position = "identity", alpha = 1/5)
ggplot(data=diamonds, mapping = aes(x = cut, colour = clarity)) +
  geom_bar(position = "identity", fill = NA)
```

* position = "fill" works like stacking, but makes each set of stacked bars the same height. This makes it easier to compare proportions across groups.

```{r}
ggplot(data = diamonds, mapping = aes(x=cut, fill = clarity)) + 
  geom_bar(position = "fill")
```

* position = "dodge" places overlapping objects directly beside one another. This makes it easier to compare individual values.

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity))+
  geom_bar(position = "dodge")
```

* Scatter plots. The values of hwy and displ are rounded so the points appear on a grid and many points overlap each other. This problem is known as overplotting. This arrangement makes it hard to see where the mass of the data is. Are the data points spread equally throughout the graph, or is there one special combination of hwy and displ that contains 109 values?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(position = "jitter")
```

### 3.8.1 Exercises
> 1. What is the problem with this plot? How could you improve it?

it may be not showing the points spread on the graph. Add position = "jitter".

_There is overplotting because there are multiple observations for each combination of cty and hwy values. I would improve the plot by using a jitter position adjustment to decrease overplotting._

_The relationship between cty and hwy is clear even without jittering the points but jittering shows the locations where there are more observations._

> 2. What parameters to geom_jitter() control the amount of jittering?

From the geom_jitter() documentation, there are two arguments to jitter:

width controls the amount of horizontal displacement, and
height controls the amount of vertical displacement.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(position = "jitter")
##Same as below
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(position = position_jitter())

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(position = position_jitter(width = 10, height = 10))
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_jitter(width = 10, height = 10)

```




> 3. Compare and contrast geom_jitter() with geom_count().

geom_jitter() uses stat = "identity", while geom_count() uses stat = "sum".
geom_jitter() uses position = "jitter" while geom_count() uses position = "identity"

_The geom geom_jitter() adds random variation to the locations points of the graph. In other words, it “jitters” the locations of points slightly. This method reduces overplotting since two points with the same location are unlikely to have the same random variation. However, the reduction in overlapping comes at the cost of slightly changing the x and y values of the points._


_The geom geom_count() sizes the points relative to the number of observations. Combinations of (x, y) values with more observations will be larger than those with fewer observations._

_The geom_count() geom does not change x and y coordinates of the points. However, if the points are close together and counts are large, the size of some points can itself create overplotting. For example, in the following example, a third variable mapped to color is added to the plot. In this case, geom_count() is less readable than geom_jitter() when adding a third variable as a color aesthetic._

```{r}
ggplot(data = mpg, aes(x = hwy, y = cty, color = drv)) +
  geom_count()

ggplot(data = mpg, aes(x = hwy, y = cty, color = drv)) +
  geom_jitter()
```

_Combining geom_count() with jitter, which is specified with the position argument to geom_count() rather than its own geom, helps overplotting a little._

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy, color = class)) +
  geom_count(position = "jitter")
```


> 4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.

  position = "dodge2"
  
_The default position for geom_boxplot() is "dodge2", which is a shortcut for position_dodge2. This position adjustment does not change the vertical position of a geom but moves the geom horizontally to avoid overlapping other geoms._

_If position_identity() is used the boxplots overlap._

```{r}
ggplot(mpg, mapping = aes(x = displ, y = hwy))+
  geom_boxplot()

ggplot(mpg, mapping = aes(x = displ, y = hwy, colour = drv))+
  geom_boxplot()
ggplot(mpg, mapping = aes(x = class, y = hwy, colour = drv))+
  geom_boxplot()

ggplot(mpg, mapping = aes(x = class, y = hwy, colour = drv))+
  geom_boxplot(position = "identity")
```


### 3.9 Coordinate systems

There are a number of other coordinate systems that are occasionally helpful.

* coord_flip() switches the x and y axes.

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy, color = class)) + geom_boxplot()

ggplot(data = mpg, mapping = aes(x = class, y = hwy, color = class)) + geom_boxplot() + coord_flip()
```

* coord_quickmap() sets the aspect ratio correctly for maps. This is very important if you’re plotting spatial data with ggplot2 (which unfortunately we don’t have the space to cover in this book).

```{r}
ug <- map_data("world")

ggplot(ug, aes(long, lat, group = group)) +
  geom_polygon(fill = "brown", colour = "red")

ggplot(ug, aes(long, lat, group = group)) +
  geom_polygon(fill = "brown", colour = "red") + coord_quickmap()

```
 

* coord_polar() uses polar coordinates. Polar coordinates reveal an interesting connection between a bar chart and a Coxcomb chart.

```{r}
bar <- ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut), show.legend = FALSE, width = 1)+
  labs(x = NULL, y = NULL)
bar
bar + coord_flip()
bar + coord_polar()

```

3.9.1 Exercises

> 1. Turn a stacked bar chart into a pie chart using coord_polar().

```{r}
stack <- ggplot(mpg, aes(x = factor(1), fill = drv)) +
  geom_bar(width = 1)
stack

stack+coord_polar()
#The argument theta = "y" maps y to the angle of each section.

stack + coord_polar(theta = "y")

```



see above

> 2. What does labs() do? Read the documentation.

It helps us modify axes, legends and plot labels such as title, subtitle, legend, caption, tags, etc.

_The labs() function is not the only function that adds titles to plots. The xlab(), ylab(), and x- and y-scale functions can add axis titles. The ggtitle() function adds plot titles._


> 3. What’s the difference between coord_quickmap() and coord_map()?

coord_quickmap is a quick approximation of a map on 2D that does preserve straight lines. coord_map requires mapproj package.

```{r}
ggplot(data = ug, mapping = aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "brown", colour = "red") + coord_quickmap()

# ggplot(data = ug, mapping = aes(x = long, y = lat, group = group)) +
#   geom_polygon(fill = "brown", colour = "red") + coord_map()
```

_The coord_map() function uses map projections to project the three-dimensional Earth onto a two-dimensional plane. By default, coord_map() uses the Mercator projection. This projection is applied to all the geoms in the plot. The coord_quickmap() function uses an approximate but faster map projection. This approximation ignores the curvature of Earth and adjusts the map for the latitude/longitude ratio. The coord_quickmap() project is faster than coord_map() both because the projection is computationally easier, and unlike coord_map(), the coordinates of the individual geoms do not need to be transformed._

> 4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() + ##It provides a reference line for a line of symmetry
  coord_fixed() ## It helps balance points on the same scale
```

_The function coord_fixed() ensures that the line produced by geom_abline() is at a 45-degree angle. A 45-degree line makes it easy to compare the highway and city mileage to the case in which city and highway MPG were equal._

_On average, humans are best able to perceive differences in angles relative to 45 degrees. See Cleveland (1993b), Cleveland (1994),Cleveland (1993a), Cleveland, McGill, and McGill (1988), Heer and Agrawala (2006) for discussion on how the aspect ratio of a plot affects perception of the values it encodes, evidence that 45-degrees is generally the optimal aspect ratio, and methods to calculate the optimal aspect ratio of a plot. The function ggthemes::bank_slopes() will calculate the optimal aspect ratio to bank slopes to 45-degrees._


## 3.10 The layered grammar of graphics


ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>


_You could use this method to build any plot that you imagine. In other words, you can use the code template that you’ve learned in this chapter to build hundreds of thousands of unique plots._






<br>

### References

Cleveland, William S. 1993a. “A Model for Studying Display Methods of Statistical Graphics.” Journal of Computational and Graphical Statistics 2 (4). Taylor & Francis: 323–43. https://doi.org/10.1080/10618600.1993.10474616.

Cleveland, William S. 1993b. Visualizing Information. Hobart Press.

Cleveland, William S. 1994. The Elements of Graphing Data. Hobart Press.

Cleveland, William S., Marylyn E. McGill, and Robert McGill. 1988. “The Shape Parameter of a Two-Variable Graph.” Journal of the American Statistical Association 83 (402). [American Statistical Association, Taylor & Francis, Ltd.]: 289–300. https://www.jstor.org/stable/2288843.

Heer, Jeffrey, and Maneesh Agrawala. 2006. “Multi-Scale Banking to 45º.” Ieee Transactions on Visualization and Computer Graphics 12 (5, September/October). https://doi.org/10.1109/TVCG.2006.163.
