---
title: "Chapter 14 to 16"
output: html_notebook
---

## 14. Strings

### 14.3 Matching patterns with regular expressions

Basic matches

```{r}
library(tidyverse)
x <- c("apples", "bananas", "pears", "man.goes")
str_view(x, ".a.")
str_view_all(x, ".a.")
str_view(x, "\\.")
str_view(x, "n\\.")
y <- "a\\b"
y; writeLines(y)
str_view(y, "\\\\")
```

*Anchors*

By default, regular expressions will match any part of a string. It’s often useful to anchor the regular expression so that it matches from the start or end of the string. You can use:

^ to match the start of the string.
$ to match the end of the string.
You can also match the boundary between words with \b "\bsum\b" is the same as "^sum$"


*14.3.3 Character classes and alternatives*

There are a number of special patterns that match more than one character. You’ve already seen ., which matches any character apart from a newline. There are four other useful tools:

\d: matches any digit.
\s: matches any whitespace (e.g. space, tab, newline).
[abc]: matches a, b, or c.
[^abc]: matches anything except a, b, or c.
Remember, to create a regular expression containing \d or \s, you’ll need to escape the \ for the string, so you’ll type "\\d" or "\\s".

This works for most (but not all) regex metacharacters: $ . | ? * + ( ) [ {. Unfortunately, a few characters have special meaning even inside a character class and must be handled with backslash escapes: ] \ ^ and -.

You can use alternation to pick between one or more alternative patterns. For example, abc|d..f will match either ‘“abc”’, or "deaf". Note that the precedence for | is low, so that abc|xyz matches abc or xyz not abcyz or abxyz. Like with mathematical expressions, if precedence ever gets confusing, use parentheses to make it clear what you want:

```{r}
str_view(c("abc", "a.c", "a*c", "a c"), "a[.]c")
str_view(c("abc", "a.c", "a*c", "a c"), ".[*]c")
str_view(c("abc", "a.c", "a*c", "a c"), "a[ ]")

str_view(c("grey", "gray"), "gr(e|a)y")

```

*14.3.4 Repetition*
The next step up in power involves controlling how many times a pattern matches:

?: 0 or 1
+: 1 or more
*: 0 or more


*14.3.5 Grouping and backreferences*
Earlier, you learned about parentheses as a way to disambiguate complex expressions. Parentheses also create a numbered capturing group (number 1, 2 etc.). A capturing group stores the part of the string matched by the part of the regular expression inside the parentheses. You can refer to the same text as previously matched by a capturing group with backreferences, like \1, \2 etc. For example, the following regular expression finds all fruits that have a repeated pair of letters.

```{r}
str_view(fruit, "(p)\\1", match = TRUE)
str_view(fruit, "(an)\\1", match = TRUE)
str_view(fruit, "(.)\\1", match = TRUE)
str_view(fruit, "(..)\\1", match = TRUE)

str_view(fruit, "(.)(.)\\2\\1", match = TRUE)
str_view(fruit, "(.).\\1.\\1", match = TRUE)
str_view(fruit, "(.)(.)(.).*\\3\\2\\1", match = TRUE)
```












