---
title: "Chapter 14 to 16"
output: html_notebook
---

## 14. Strings

### 14.3 Matching patterns with regular expressions

Basic matches

```{r}
library(tidyverse)
x <- c("apples", "bananas", "pears", "man.goes")
str_view(x, ".a.")
str_view_all(x, ".a.")
str_view(x, "\\.")
str_view(x, "n\\.")
y <- "a\\b"
y; writeLines(y)
str_view(y, "\\\\")
```

*Anchors*

By default, regular expressions will match any part of a string. It’s often useful to anchor the regular expression so that it matches from the start or end of the string. You can use:

^ to match the start of the string.
$ to match the end of the string.
You can also match the boundary between words with \b "\bsum\b" is the same as "^sum$"


*14.3.3 Character classes and alternatives*

There are a number of special patterns that match more than one character. You’ve already seen ., which matches any character apart from a newline. There are four other useful tools:

\d: matches any digit.
\s: matches any whitespace (e.g. space, tab, newline).
[abc]: matches a, b, or c.
[^abc]: matches anything except a, b, or c.
Remember, to create a regular expression containing \d or \s, you’ll need to escape the \ for the string, so you’ll type "\\d" or "\\s".

This works for most (but not all) regex metacharacters: $ . | ? * + ( ) [ {. Unfortunately, a few characters have special meaning even inside a character class and must be handled with backslash escapes: ] \ ^ and -.

You can use alternation to pick between one or more alternative patterns. For example, abc|d..f will match either ‘“abc”’, or "deaf". Note that the precedence for | is low, so that abc|xyz matches abc or xyz not abcyz or abxyz. Like with mathematical expressions, if precedence ever gets confusing, use parentheses to make it clear what you want:

```{r}
str_view(c("abc", "a.c", "a*c", "a c"), "a[.]c")
str_view(c("abc", "a.c", "a*c", "a c"), ".[*]c")
str_view(c("abc", "a.c", "a*c", "a c"), "a[ ]")

str_view(c("grey", "gray"), "gr(e|a)y")

```

*14.3.4 Repetition*
The next step up in power involves controlling how many times a pattern matches:

?: 0 or 1
+: 1 or more
*: 0 or more


*14.3.5 Grouping and backreferences*
Earlier, you learned about parentheses as a way to disambiguate complex expressions. Parentheses also create a numbered capturing group (number 1, 2 etc.). A capturing group stores the part of the string matched by the part of the regular expression inside the parentheses. You can refer to the same text as previously matched by a capturing group with backreferences, like \1, \2 etc. For example, the following regular expression finds all fruits that have a repeated pair of letters.

```{r}
str_view(fruit, "(p)\\1", match = TRUE)
str_view(fruit, "(an)\\1", match = TRUE)
str_view(fruit, "(.)\\1", match = TRUE)
str_view(fruit, "(..)\\1", match = TRUE)

str_view(fruit, "(.)(.)\\2\\1", match = TRUE)
str_view(fruit, "(.).\\1.\\1", match = TRUE)
str_view(fruit, "(.)(.)(.).*\\3\\2\\1", match = TRUE)
```


### 14.4 Tools

In this section you’ll learn a wide array of stringr functions that let you:

* Determine which strings match a pattern.
* Find the positions of matches.
* Extract the content of matches.
* Replace matches with new values.
* Split a string based on a match.

```{r}
x <- c("apple", "banana", "apple", "oranges")

str_detect(x, "e") ##returns a logical vector
#hwo many words in words dataset end with d
sum(str_detect(words, "d$"))
#proportion that starts with a vowel, vowels are aeiou. How about ends wit a vowel
mean(str_detect(words, "^[aeiou]")) #start
mean(str_detect(words, "[aeiou]$")) #ends

# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
words[247:251]
str_detect(words[247:251], "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)

#A common use of str_detect() is to select the elements that match a pattern.
#select words that end with x
words[str_detect(words, "x$")]
##or just
str_subset(words, "x$")
```


```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")

mean(str_count(words, "[aeiou]"))

str_view(x, "a"); str_view_all(x, "a")

head(sentences)
length(sentences)

colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match

has_colour <- str_subset(sentences, colour_match)
head(has_colour)
length(has_colour)
matches <- str_extract(has_colour, colour_match)
head(matches)
length(matches)
more <- sentences[str_count(sentences, colour_match)>1]
head(more)
str_view_all(more, colour_match)
str_extract(more, colour_match) #first matches are selected per sentence
str_extract_all(more, colour_match) #returns a list with all matches
str_extract_all(more, colour_match, simplify = TRUE)
##more extract examples

x<- c("a", "ab", "a bc", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

#### 14.4.3 Grouped Matches

```{r}
noun <- "(a|the) ([^ ]+)"
(has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10))
has_noun %>% str_extract(noun)
#str_match() gives each individual component. Instead of a character vector, it returns a matrix, with one column for the complete match followed by one column for each group:
has_noun %>% str_match(noun)

tibble(sentence=sentences) %>%
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", remove = FALSE
  )

```

#### replacing matches
```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
str_replace_all(x, "[aeiou]", "-")

#Instead of replacing with a fixed string you can use backreferences to insert components of the match. In the following code, I flip the order of the second and third words.
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

#### 14.4.5 Splitting
```{r}
sentences %>% head(5) %>% str_split(" ")
"a|b|c|d" %>% str_split("\\|") %>% .[[1]]
sentences %>% head(5) %>% str_split(" ", simplify = TRUE)
##very useful in seperating key value pairs
fields <- c("First name: Wilbrod", "Surname: Ntawiha", "Country: UG", "Age: 31")
fields %>% str_split(": ", n = 2, simplify = TRUE)

#Instead of splitting up strings by patterns, you can also split up by character, line, sentence and word boundary()s
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
str_split(x, " ")[[1]]
str_split(x, boundary("word"))[[1]]

##MATCHES
str_locate(x, "is")
str_locate_all(x, "is")
```

#### other ideas

```{r}
fruit <- c("apple", "bananas", "BANANAS", "Bananas")
str_view(fruit, "banana") # is a short form of below
str_view(fruit, regex("banana"))
str_view(fruit, regex("banana", ignore_case = T))

#multiline = TRUE allows ^ and $ to match the start and end of each line rather than the start and end of the complete string.
x <- "Line 1\nLine 2\nLine 3"
str_extract_all(x, "^Line")[[1]]
str_extract_all(x, regex("^Line", multiline = TRUE))[[1]]

#comments = TRUE allows you to use comments and white space to make complex regular expressions more understandable. Spaces are ignored, as is everything after #. To match a literal space, you’ll need to escape it: "\\ ".
phone <- regex("
  \\(?     # optional opening parens
  (\\d{3}) # area code
  [) -]?   # optional closing parens, space, or dash
  (\\d{3}) # another three numbers
  [ -]?    # optional space or dash
  (\\d{3}) # three more numbers
  ", comments = TRUE)

str_match("514-791-8141", phone)

#dotall = TRUE allows . to match everything, including \n.

#fixed(): matches exactly the specified sequence of bytes. It ignores all special regular expressions and operates at a very low level. This allows you to avoid complex escaping and can be much faster than regular expressions. The following microbenchmark shows that it’s about 3x faster for a simple example.
microbenchmark::microbenchmark(
  fixed = str_detect(sentences, fixed("the")),##useful for english matches without special characters
  regex = str_detect(sentences, "the"),
  times = 20
)
# you can use coll() instead of fixed() for special cases. coll(): compare strings using standard collation rules.

```


```{r}
# That means you also need to be aware of the difference
# when doing case insensitive matches:
i <- c("I", "İ", "i", "ı")
i
str_subset(i, coll("i", ignore_case = TRUE))
str_subset(i, coll("i", ignore_case = TRUE, locale = "tr"))

stringi::stri_locale_info()

##Other uses of regular expressions
apropos("replace") #apropos() searches all objects available from the global environment. This is useful if you can’t quite remember the name of the function.
#dir() lists all the files in a directory. The pattern argument takes a regular expression and only returns file names that match the pattern. For example, you can find all the R Markdown files in the current directory with:
head(dir(pattern = "\\.Rmd$"))

```


# Factors

### read the book and re check ordering factors by x axis in ggplot charts


# Dates and Time

```{r}
library(lubridate)
today()
now()

ymd("2020-09-20")
mdy("September 20th, 2020")
dmy("20/9/2020")
ymd("20200930", tz = "UTC")

ymd_hms("20201030 11:27:15")

##How old is wilbrod?
w_age <- today() - ymd(19890621)
w_age
as.duration(w_age)

##intervals
years(1)/days(1) #supposed to be 366
next_year <- today() + years(1)
(today() %--% next_year) / ddays(1)
(today() %--% next_year) %/% days(1) ##integer division much better

##Timezones
head(OlsonNames())
t1 <- now()
t11 <- with_tz(t1, tzone = "Australia/Lord_Howe")
t111 <- force_tz(t1, tzone = "Australia/Lord_Howe")

t1; t11; t111
```












