---
title: "R Notebook"
output: html_notebook
---

# 4 Workflow: basics

_Frustration is natural when you start programming in R, because it is such a stickler for punctuation, and even one character out of place will cause it to complain. But while you should expect to be a little frustrated, take comfort in that it’s both typical and temporary: it happens to everyone, and the only way to get over it is to keep trying._

### Basics

* You can use R as a calculator:
* You can create new objects with <-

_All R statements where you create objects, assignment statements, have the same form: object_name <- value. When reading that code say “object name gets value” in your head. You will make lots of assignments and <- is a pain to type. Don’t be lazy and use =, it will work, but it will cause confusion later. Instead, use RStudio’s keyboard shortcut: Alt + - (the minus sign). Notice that RStudio automagically surrounds <- with spaces, which is a good code formatting practice. Code is miserable to read on a good day, so giveyoureyesabreak and use spaces._


### What’s in a name?

Object names must start with a letter, and can only contain letters, numbers, _ and .. You want your object names to be descriptive, so you’ll need a convention for multiple words. I recommend snake_case where you separate lowercase words with _.

* i_use_snake_case
* otherPeopleUseCamelCase
* some.people.use.periods
* And_aFew.People_RENOUNCEconvention

**Typos matter. Case matters.**

##  Calling functions

```{r}
(x <- seq(1, to = 10, by = 2))
```

* Quotation marks and parentheses must always come in a pair.

### 4.4 Exercises

1. variable names are different

2. 

```{r}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)
filter(diamonds, carat > 3)
```


3. Tools -> Key board shortcuts help (Alt + Shift + K)








