---
title: "Chapter 17:"
output: html_notebook
---

# Pipes
## Introduction

2 more books to cover include:

 * Hands on Programming with R, by Garrett Grolemund.
 * Advanced R by Hadley Wickham: http://adv-r.had.co.nz
 
The pipe, %>%, comes from the magrittr package by Stefan Milton Bache. Packages in the tidyverse load %>% for you automatically, so you don’t usually load magrittr explicitly. Here, however, we’re focussing on piping, and we aren’t loading any other packages, so we will load it explicitly.

```{r}
library(magrittr)

```

 
## Piping alternatives

Piping makes code easier to read and understand

## 21. Iteration

```{r}
seq(0, 1, length.out = 11)
seq(stats::rnorm(20)) # effectively 'along'
seq(1, 9, by = 2)     # matches 'end'
seq(1, 9, by = pi)    # stays below 'end'
seq(1, 6, by = 3)
seq(1.575, 5.125, by = 0.05)
seq(17) # same as 1:17, or even better seq_len(17)

means <- c(0, 1, 4)

seq_along(means)
```

```{r}
x = 100
x
for (i in seq(2)) {
  x=x*i
  x
}
x

```

### map() functions

map() makes a list.
map_lgl() makes a logical vector.
map_int() makes an integer vector.
map_dbl() makes a double vector.
map_chr() makes a character vector.

```{r}
library(tidyverse)
mtcars %>% split(.$cyl) %>%
  map(function(df){lm(mpg ~ wt, data = df)})

## alternatively using .f

mtcars%>% split(.$cyl) %>%
  map(~lm(mpg~wt, data = .))

##to obtain a summary result 
mtcars %>% split(.$cyl) %>%
  map(~lm(mpg~wt, data = .))%>%
  map(summary)%>%
  map_dbl("r.squared") 
##or
mtcars %>% split(.$cyl) %>%
  map(~lm(mpg~wt, data = .))%>%
  map(summary)%>%
  map_dbl(~.$r.squared)
```

```{r}
library(tidyverse)
x = list(1, "liz", 8.989)
print(x)

walk(x,print)

x%>%walk(print)

mtcars
mtcars %>% 
  split(.$cyl) %>% 
  map(~ggplot(., aes(mpg, wt)) + geom_point()) -> plots
plots
paths <- str_c("plot", names(plots), ".pdf")
#paths <- paste("plot", names(plots), ".pdf")
paths

pwalk(list(paths, plots), ggsave, path = getwd())




```





